---
title: "prepare and vis ADMIXTURE"
author: "Devon DeRaad"
date: "3/8/2022"
output: html_document
---

```{r}
library(vcfR)
library(SNPfiltR)
library(ggplot2)
```

### Filter input SNP files based on MAC to increase signal to noise ratio
```{r}
#bring in sample info
#read in sample info csv
sample.info<-read.csv("~/Desktop/mex.towhees/sample.locs.csv")

#filter input vcf by MAC in order to increase signal to noise ratio for ADMIXTURE analysis
#read in filtered vcf
vcfR <- read.vcfR("~/Desktop/mex.towhees/towhee.75.mac2.unlinked.nomito.vcf")
vcfR

#filter by MAC thresholds
vcf.3<-min_mac(vcfR, min.mac = 3)
vcf.3

#filter by MAC thresholds
vcf.4<-min_mac(vcfR, min.mac = 4)
vcf.4

#filter by MAC thresholds
vcf.5<-min_mac(vcfR, min.mac = 5)
vcf.5

#filter by MAC thresholds
vcf.6<-min_mac(vcfR, min.mac = 6)
vcf.6

#filter by MAC thresholds
vcf.7<-min_mac(vcfR, min.mac = 7)
vcf.7

#filter by MAC thresholds
vcf.8<-min_mac(vcfR, min.mac = 8)
vcf.8

#filter by MAC thresholds
vcf.9<-min_mac(vcfR, min.mac = 9)
vcf.9

#filter by MAC thresholds
vcf.10<-min_mac(vcfR, min.mac = 10)
vcf.10

#write out vcf file
#vcfR::write.vcf(vcf.3, file="~/Downloads/towhee.mac3.vcf.gz")
#vcfR::write.vcf(vcf.4, file="~/Downloads/towhee.mac4.vcf.gz")
#vcfR::write.vcf(vcf.5, file="~/Downloads/towhee.mac5.vcf.gz")
#vcfR::write.vcf(vcf.6, file="~/Downloads/towhee.mac6.vcf.gz")
#vcfR::write.vcf(vcf.7, file="~/Downloads/towhee.mac7.vcf.gz")
#vcfR::write.vcf(vcf.8, file="~/Downloads/towhee.mac8.vcf.gz")
#vcfR::write.vcf(vcf.9, file="~/Downloads/towhee.mac9.vcf.gz")
#vcfR::write.vcf(vcf.10, file="~/Downloads/towhee.mac10.vcf.gz")

```

### Visualize ADMIXTURE results for MAC 2
```{r}
#setwd to admixture directory run on the cluster
setwd("~/Desktop/mex.towhees/admixture.mac2/")
#read in log error values to determine optimal K
log<-read.table("log.errors.txt")[,c(3:4)]
#use double backslash to interpret the opening parentheses literally in the regular expression
log$V3<-gsub("\\(K=", "", log$V3)
log$V3<-gsub("):", "", log$V3)
#interpret K values as numerical
log$V3<-as.numeric(log$V3)
#rename columns
colnames(log)<-c("Kvalue","cross.validation.error")
#make plot showing the cross validation error across K values 1:10
ggplot(data=log, aes(x=Kvalue, y=cross.validation.error, group=1)) +
  geom_line(linetype = "dashed")+
  geom_point()+
  ylab("cross-validation error")+
  xlab("K")+
  scale_x_continuous(breaks = c(1:5))+
  theme_classic()
```

```{r, fig.height=10, fig.width=8}
#setwd to admixture directory run on the cluster
setwd("~/Desktop/mex.towhees/admixture.mac2/")
#read in input file in order to get list of input samples in order
samps<-read.table("binary_fileset.fam")[,1]

#reorder sampling df to match order of the plot
sample.info<-sample.info[match(samps, sample.info$ID),]
sample.info$ID == samps

#read in all ten runs and save each dataframe in a list
runs<-list()
#read in log files
for (i in 1:5){
  runs[[i]]<-read.table(paste0("binary_fileset.", i, ".Q"))
}
par(mfrow=c(5,1))
#plot each run
for (i in 1:5){
barplot(t(as.matrix(runs[[i]])), col=rainbow(i), ylab="Ancestry", border="black")
}

#set rownames
rownames(sample.info)<-c(1:32)
#get brewer cols
RColorBrewer::brewer.pal(n=8,"RdGy")
#reorder samples based on sampling locality for k=2/3
runs[[2]]<-runs[[2]][c(7:10,32,29,30,31,3:6,23,24,21,22,27,28,25,26,11:14,15:18,2,1,19,20),]
runs[[3]]<-runs[[3]][c(7:10,32,29,30,31,3:6,23,24,21,22,27,28,25,26,11:14,15:18,2,1,19,20),]
#plot barplots for the two most relevant runs
barplot(t(as.matrix(runs[[2]])), col=c("#B2182B","#4D4D4D"), ylab="Ancestry", border="black")
barplot(t(as.matrix(runs[[3]])), col=c("#B2182B","#4D4D4D","black"), ylab="Ancestry", border="black")
```


### Visualize ADMIXTURE results for MAC 3
```{r}
#setwd to admixture directory run on the cluster
setwd("~/Desktop/mex.towhees/admixture.mac3/")
#read in log error values to determine optimal K
log<-read.table("log.errors.txt")[,c(3:4)]
#use double backslash to interpret the opening parentheses literally in the regular expression
log$V3<-gsub("\\(K=", "", log$V3)
log$V3<-gsub("):", "", log$V3)
#interpret K values as numerical
log$V3<-as.numeric(log$V3)
#rename columns
colnames(log)<-c("Kvalue","cross.validation.error")
#make plot showing the cross validation error across K values 1:10
ggplot(data=log, aes(x=Kvalue, y=cross.validation.error, group=1)) +
  geom_line(linetype = "dashed")+
  geom_point()+
  ylab("cross-validation error")+
  xlab("K")+
  scale_x_continuous(breaks = c(1:10))+
  theme_classic()
```

```{r, fig.height=10, fig.width=8}
#setwd to admixture directory run on the cluster
setwd("~/Desktop/mex.towhees/admixture.mac3/")
#read in all ten runs and save each dataframe in a list
runs<-list()
#read in log files
for (i in 1:5){
  runs[[i]]<-read.table(paste0("binary_fileset.", i, ".Q"))
}
par(mfrow=c(5,1))
#plot each run
for (i in 1:5){
barplot(t(as.matrix(runs[[i]])), col=rainbow(i), ylab="Ancestry", border="black")
}

#set rownames
rownames(sample.info)<-c(1:32)
#get brewer cols
RColorBrewer::brewer.pal(n=8,"RdGy")
#reorder samples based on sampling locality for k=2/3
runs[[2]]<-runs[[2]][c(7:10,32,29,30,31,3:6,23,24,21,22,27,28,25,26,11:14,15:18,2,1,19,20),]
runs[[3]]<-runs[[3]][c(7:10,32,29,30,31,3:6,23,24,21,22,27,28,25,26,11:14,15:18,2,1,19,20),]
#plot barplots for the two most relevant runs
barplot(t(as.matrix(runs[[2]])), col=c("#B2182B","#4D4D4D"), ylab="Ancestry", border="black")
barplot(t(as.matrix(runs[[3]])), col=c("#B2182B","#4D4D4D","black"), ylab="Ancestry", border="black")
```


### Visualize ADMIXTURE results for MAC 4
```{r}
#setwd to admixture directory run on the cluster
setwd("~/Desktop/mex.towhees/admixture.mac4/")
#read in log error values to determine optimal K
log<-read.table("log.errors.txt")[,c(3:4)]
#use double backslash to interpret the opening parentheses literally in the regular expression
log$V3<-gsub("\\(K=", "", log$V3)
log$V3<-gsub("):", "", log$V3)
#interpret K values as numerical
log$V3<-as.numeric(log$V3)
#rename columns
colnames(log)<-c("Kvalue","cross.validation.error")
#make plot showing the cross validation error across K values 1:10
ggplot(data=log, aes(x=Kvalue, y=cross.validation.error, group=1)) +
  geom_line(linetype = "dashed")+
  geom_point()+
  ylab("cross-validation error")+
  xlab("K")+
  scale_x_continuous(breaks = c(1:10))+
  theme_classic()
```

```{r, fig.height=10, fig.width=8}
#setwd to admixture directory run on the cluster
setwd("~/Desktop/mex.towhees/admixture.mac4/")
#read in all ten runs and save each dataframe in a list
runs<-list()
#read in log files
for (i in 1:5){
  runs[[i]]<-read.table(paste0("binary_fileset.", i, ".Q"))
}
par(mfrow=c(5,1))
#plot each run
for (i in 1:5){
barplot(t(as.matrix(runs[[i]])), col=rainbow(i), ylab="Ancestry", border="black")
}

#set rownames
rownames(sample.info)<-c(1:32)
#get brewer cols
RColorBrewer::brewer.pal(n=8,"RdGy")
#reorder samples based on sampling locality for k=2/3
runs[[2]]<-runs[[2]][c(7:10,32,29,30,31,3:6,23,24,21,22,27,28,25,26,11:14,15:18,2,1,19,20),]
runs[[3]]<-runs[[3]][c(7:10,32,29,30,31,3:6,23,24,21,22,27,28,25,26,11:14,15:18,2,1,19,20),]
#plot barplots for the two most relevant runs
barplot(t(as.matrix(runs[[2]])), col=c("#B2182B","#4D4D4D"), ylab="Ancestry", border="black")
barplot(t(as.matrix(runs[[3]])), col=c("#B2182B","#4D4D4D","black"), ylab="Ancestry", border="black")
```


### Visualize ADMIXTURE results for MAC 5
```{r}
#setwd to admixture directory run on the cluster
setwd("~/Desktop/mex.towhees/admixture.mac5/")
#read in log error values to determine optimal K
log<-read.table("log.errors.txt")[,c(3:4)]
#use double backslash to interpret the opening parentheses literally in the regular expression
log$V3<-gsub("\\(K=", "", log$V3)
log$V3<-gsub("):", "", log$V3)
#interpret K values as numerical
log$V3<-as.numeric(log$V3)
#rename columns
colnames(log)<-c("Kvalue","cross.validation.error")
#make plot showing the cross validation error across K values 1:10
ggplot(data=log, aes(x=Kvalue, y=cross.validation.error, group=1)) +
  geom_line(linetype = "dashed")+
  geom_point()+
  ylab("cross-validation error")+
  xlab("K")+
  scale_x_continuous(breaks = c(1:10))+
  theme_classic()
```

```{r, fig.height=10, fig.width=8}
#setwd to admixture directory run on the cluster
setwd("~/Desktop/mex.towhees/admixture.mac5/")
#read in all ten runs and save each dataframe in a list
runs<-list()
#read in log files
for (i in 1:5){
  runs[[i]]<-read.table(paste0("binary_fileset.", i, ".Q"))
}
par(mfrow=c(5,1))
#plot each run
for (i in 1:5){
barplot(t(as.matrix(runs[[i]])), col=rainbow(i), ylab="Ancestry", border="black")
}

#set rownames
rownames(sample.info)<-c(1:32)
#get brewer cols
RColorBrewer::brewer.pal(n=8,"RdGy")
#reorder samples based on sampling locality for k=2/3
runs[[2]]<-runs[[2]][c(7:10,32,29,30,31,3:6,23,24,21,22,27,28,25,26,11:14,15:18,2,1,19,20),]
runs[[3]]<-runs[[3]][c(7:10,32,29,30,31,3:6,23,24,21,22,27,28,25,26,11:14,15:18,2,1,19,20),]
#plot barplots for the two most relevant runs
barplot(t(as.matrix(runs[[2]])), col=c("#B2182B","#4D4D4D"), ylab="Ancestry", border="black")
barplot(t(as.matrix(runs[[3]])), col=c("#B2182B","#4D4D4D","black"), ylab="Ancestry", border="black")
```


### Visualize ADMIXTURE results for MAC 6
```{r}
#setwd to admixture directory run on the cluster
setwd("~/Desktop/mex.towhees/admixture.mac6/")
#read in log error values to determine optimal K
log<-read.table("log.errors.txt")[,c(3:4)]
#use double backslash to interpret the opening parentheses literally in the regular expression
log$V3<-gsub("\\(K=", "", log$V3)
log$V3<-gsub("):", "", log$V3)
#interpret K values as numerical
log$V3<-as.numeric(log$V3)
#rename columns
colnames(log)<-c("Kvalue","cross.validation.error")
#make plot showing the cross validation error across K values 1:10
ggplot(data=log, aes(x=Kvalue, y=cross.validation.error, group=1)) +
  geom_line(linetype = "dashed")+
  geom_point()+
  ylab("cross-validation error")+
  xlab("K")+
  scale_x_continuous(breaks = c(1:10))+
  theme_classic()
```

```{r, fig.height=10, fig.width=8}
#setwd to admixture directory run on the cluster
setwd("~/Desktop/mex.towhees/admixture.mac6/")
#read in all ten runs and save each dataframe in a list
runs<-list()
#read in log files
for (i in 1:5){
  runs[[i]]<-read.table(paste0("binary_fileset.", i, ".Q"))
}
par(mfrow=c(5,1))
#plot each run
for (i in 1:5){
barplot(t(as.matrix(runs[[i]])), col=rainbow(i), ylab="Ancestry", border="black")
}

#set rownames
rownames(sample.info)<-c(1:32)
#get brewer cols
RColorBrewer::brewer.pal(n=8,"RdGy")
#reorder samples based on sampling locality for k=2/3
runs[[2]]<-runs[[2]][c(7:10,32,29,30,31,3:6,23,24,21,22,27,28,25,26,11:14,15:18,2,1,19,20),]
runs[[3]]<-runs[[3]][c(7:10,32,29,30,31,3:6,23,24,21,22,27,28,25,26,11:14,15:18,2,1,19,20),]
#plot barplots for the two most relevant runs
barplot(t(as.matrix(runs[[2]])), col=c("#B2182B","#4D4D4D"), ylab="Ancestry", border="black")
barplot(t(as.matrix(runs[[3]])), col=c("#B2182B","#4D4D4D","black"), ylab="Ancestry", border="black")
```


### Visualize ADMIXTURE results for MAC 7
```{r}
#setwd to admixture directory run on the cluster
setwd("~/Desktop/mex.towhees/admixture.mac7/")
#read in log error values to determine optimal K
log<-read.table("log.errors.txt")[,c(3:4)]
#use double backslash to interpret the opening parentheses literally in the regular expression
log$V3<-gsub("\\(K=", "", log$V3)
log$V3<-gsub("):", "", log$V3)
#interpret K values as numerical
log$V3<-as.numeric(log$V3)
#rename columns
colnames(log)<-c("Kvalue","cross.validation.error")
#make plot showing the cross validation error across K values 1:10
ggplot(data=log, aes(x=Kvalue, y=cross.validation.error, group=1)) +
  geom_line(linetype = "dashed")+
  geom_point()+
  ylab("cross-validation error")+
  xlab("K")+
  scale_x_continuous(breaks = c(1:10))+
  theme_classic()
```

```{r, fig.height=10, fig.width=8}
#setwd to admixture directory run on the cluster
setwd("~/Desktop/mex.towhees/admixture.mac7/")
#read in all ten runs and save each dataframe in a list
runs<-list()
#read in log files
for (i in 1:5){
  runs[[i]]<-read.table(paste0("binary_fileset.", i, ".Q"))
}
par(mfrow=c(5,1))
#plot each run
for (i in 1:5){
barplot(t(as.matrix(runs[[i]])), col=rainbow(i), ylab="Ancestry", border="black")
}

#set rownames
rownames(sample.info)<-c(1:32)
#get brewer cols
RColorBrewer::brewer.pal(n=8,"RdGy")
#reorder samples based on sampling locality for k=2/3
runs[[2]]<-runs[[2]][c(7:10,32,29,30,31,3:6,23,24,21,22,27,28,25,26,11:14,15:18,2,1,19,20),]
runs[[3]]<-runs[[3]][c(7:10,32,29,30,31,3:6,23,24,21,22,27,28,25,26,11:14,15:18,2,1,19,20),]
#plot barplots for the two most relevant runs
barplot(t(as.matrix(runs[[2]])), col=c("#B2182B","#4D4D4D"), ylab="Ancestry", border="black")
barplot(t(as.matrix(runs[[3]])), col=c("#B2182B","#4D4D4D","black"), ylab="Ancestry", border="black")
```


### Visualize ADMIXTURE results for MAC 8
```{r}
#setwd to admixture directory run on the cluster
setwd("~/Desktop/mex.towhees/admixture.mac8/")
#read in log error values to determine optimal K
log<-read.table("log.errors.txt")[,c(3:4)]
#use double backslash to interpret the opening parentheses literally in the regular expression
log$V3<-gsub("\\(K=", "", log$V3)
log$V3<-gsub("):", "", log$V3)
#interpret K values as numerical
log$V3<-as.numeric(log$V3)
#rename columns
colnames(log)<-c("Kvalue","cross.validation.error")
#make plot showing the cross validation error across K values 1:10
ggplot(data=log, aes(x=Kvalue, y=cross.validation.error, group=1)) +
  geom_line(linetype = "dashed")+
  geom_point()+
  ylab("cross-validation error")+
  xlab("K")+
  scale_x_continuous(breaks = c(1:10))+
  theme_classic()
```

```{r, fig.height=10, fig.width=8}
#setwd to admixture directory run on the cluster
setwd("~/Desktop/mex.towhees/admixture.mac8/")
#read in all ten runs and save each dataframe in a list
runs<-list()
#read in log files
for (i in 1:5){
  runs[[i]]<-read.table(paste0("binary_fileset.", i, ".Q"))
}
par(mfrow=c(5,1))
#plot each run
for (i in 1:5){
barplot(t(as.matrix(runs[[i]])), col=rainbow(i), ylab="Ancestry", border="black")
}

#set rownames
rownames(sample.info)<-c(1:32)
#get brewer cols
RColorBrewer::brewer.pal(n=8,"RdGy")
#reorder samples based on sampling locality for k=2/3
runs[[2]]<-runs[[2]][c(7:10,32,29,30,31,3:6,23,24,21,22,27,28,25,26,11:14,15:18,2,1,19,20),]
runs[[3]]<-runs[[3]][c(7:10,32,29,30,31,3:6,23,24,21,22,27,28,25,26,11:14,15:18,2,1,19,20),]
#plot barplots for the two most relevant runs
barplot(t(as.matrix(runs[[2]])), col=c("#B2182B","#4D4D4D"), ylab="Ancestry", border="black")
barplot(t(as.matrix(runs[[3]])), col=c("#B2182B","#4D4D4D","black"), ylab="Ancestry", border="black")
```

### Visualize ADMIXTURE results for MAC 9
```{r}
#setwd to admixture directory run on the cluster
setwd("~/Desktop/mex.towhees/admixture.mac9/")
#read in log error values to determine optimal K
log<-read.table("log.errors.txt")[,c(3:4)]
#use double backslash to interpret the opening parentheses literally in the regular expression
log$V3<-gsub("\\(K=", "", log$V3)
log$V3<-gsub("):", "", log$V3)
#interpret K values as numerical
log$V3<-as.numeric(log$V3)
#rename columns
colnames(log)<-c("Kvalue","cross.validation.error")
#make plot showing the cross validation error across K values 1:10
ggplot(data=log, aes(x=Kvalue, y=cross.validation.error, group=1)) +
  geom_line(linetype = "dashed")+
  geom_point()+
  ylab("cross-validation error")+
  xlab("K")+
  scale_x_continuous(breaks = c(1:10))+
  theme_classic()
```

```{r, fig.height=10, fig.width=8}
#setwd to admixture directory run on the cluster
setwd("~/Desktop/mex.towhees/admixture.mac9/")
#read in all ten runs and save each dataframe in a list
runs<-list()
#read in log files
for (i in 1:5){
  runs[[i]]<-read.table(paste0("binary_fileset.", i, ".Q"))
}
par(mfrow=c(5,1))
#plot each run
for (i in 1:5){
barplot(t(as.matrix(runs[[i]])), col=rainbow(i), ylab="Ancestry", border="black")
}

#set rownames
rownames(sample.info)<-c(1:32)
#get brewer cols
RColorBrewer::brewer.pal(n=8,"RdGy")
#reorder samples based on sampling locality for k=2/3
runs[[2]]<-runs[[2]][c(7:10,32,29,30,31,3:6,23,24,21,22,27,28,25,26,11:14,15:18,2,1,19,20),]
runs[[3]]<-runs[[3]][c(7:10,32,29,30,31,3:6,23,24,21,22,27,28,25,26,11:14,15:18,2,1,19,20),]
#plot barplots for the two most relevant runs
barplot(t(as.matrix(runs[[2]])), col=c("#B2182B","#4D4D4D"), ylab="Ancestry", border="black")
barplot(t(as.matrix(runs[[3]])), col=c("#B2182B","#4D4D4D","black"), ylab="Ancestry", border="black")
```

### Visualize ADMIXTURE results for MAC 10
```{r}
#setwd to admixture directory run on the cluster
setwd("~/Desktop/mex.towhees/admixture.mac10/")
#read in log error values to determine optimal K
log<-read.table("log.errors.txt")[,c(3:4)]
#use double backslash to interpret the opening parentheses literally in the regular expression
log$V3<-gsub("\\(K=", "", log$V3)
log$V3<-gsub("):", "", log$V3)
#interpret K values as numerical
log$V3<-as.numeric(log$V3)
#rename columns
colnames(log)<-c("Kvalue","cross.validation.error")
#make plot showing the cross validation error across K values 1:10
ggplot(data=log, aes(x=Kvalue, y=cross.validation.error, group=1)) +
  geom_line(linetype = "dashed")+
  geom_point()+
  ylab("cross-validation error")+
  xlab("K")+
  scale_x_continuous(breaks = c(1:10))+
  theme_classic()
```

```{r, fig.height=10, fig.width=8}
#setwd to admixture directory run on the cluster
setwd("~/Desktop/mex.towhees/admixture.mac10/")
#read in all ten runs and save each dataframe in a list
runs<-list()
#read in log files
for (i in 1:5){
  runs[[i]]<-read.table(paste0("binary_fileset.", i, ".Q"))
}
par(mfrow=c(5,1))
#plot each run
for (i in 1:5){
barplot(t(as.matrix(runs[[i]])), col=rainbow(i), ylab="Ancestry", border="black")
}

#set rownames
rownames(sample.info)<-c(1:32)
#get brewer cols
RColorBrewer::brewer.pal(n=8,"RdGy")
#reorder samples based on sampling locality for k=2/3
runs[[2]]<-runs[[2]][c(7:10,32,29,30,31,3:6,23,24,21,22,27,28,25,26,11:14,15:18,2,1,19,20),]
runs[[3]]<-runs[[3]][c(7:10,32,29,30,31,3:6,23,24,21,22,27,28,25,26,11:14,15:18,2,1,19,20),]
#plot barplots for the two most relevant runs
barplot(t(as.matrix(runs[[2]])), col=c("#B2182B","#4D4D4D"), ylab="Ancestry", border="black")
barplot(t(as.matrix(runs[[3]])), col=c("#B2182B","#4D4D4D","black"), ylab="Ancestry", border="black")
```

### Reorganize barplots to match geography of the transect
```{r}
#setwd to admixture directory run on the cluster
setwd("~/Desktop/mex.towhees/admixture.mac10/")

#save barplots
#pdf("~/Desktop/mex.towhees/admixture/admix.plots.pdf", width = 8, height=3.3)
#par(mfrow=c(2,1))
#par(mar = c(3, 3, 0, 0), oma = c(1, 1, 1, 1)) #set margins
#barplot(t(as.matrix(runs[[2]])), col=c("#B2182B","#4D4D4D"), ylab="Ancestry", border="black")
#barplot(t(as.matrix(runs[[3]])), col=c("#B2182B","#4D4D4D","black"), ylab="Ancestry", border="black")
#dev.off()
```

